Peer-to-Peer 6502 Cellular Automata

### 2018-07-06 paper.tex
labels

% Memory address
\newcommand\code[1]{{\tt #1}}
\newcommand\hex[1]{{\tt #1}}
\newcommand\hexrange[2]{\hex{#1}{\tt -}\hex{#2}}

@@

% Tables, macros
% Macro to generate neighborhood memory map table
\newcommand\memtable[1]{
\begin{tabular}{lll}
  \hline
  From & To & Contents \\
  \hline
  #1
  \hline
\end{tabular}
}

\newcommand\memrow[4]{
    \hex{{#1}00} & \hex{{#1}FF} & State of cell $(x#2,y#3,z#4)$ \\
}

% The \memxyz macro gives a densely-connected (von Neumann) neighborhood
\newcommand\memz[5]{
  \memrow{#1}{#4}{#5}{-1}
  \memrow{#2}{#4}{#5}{}
  \memrow{#3}{#4}{#5}{+1}
}

\newcommand\memyz[2]{
  \memz{{#1}0}{{#1}1}{{#1}2}{#2}{-1}
  \memz{{#1}4}{{#1}5}{{#1}6}{#2}{}
  \memz{{#1}8}{{#1}9}{{#1}A}{#2}{+1}
}

\newcommand\memxyz{
  \memyz{4}{-1}

  %  \memyz{5}{} includes zero page-mapped cell (x,y,z), which we want to skip, so spell it out without that
  \memz{50}{51}{52}{}{-1}
  \memrow{54}{}{}{-1}
  \memrow{56}{}{}{+1}
  \memz{58}{59}{5A}{}{+1}

  \memyz{6}{+1}
}

% The following table is a sparse neighborhood
\newcommand\vonneumannmap{\memtable{
  \memrow{45}{-1}{}{}
  \memrow{51}{}{-1}{}
  \memrow{54}{}{}{-1}
  \memrow{56}{}{}{+1}
  \memrow{59}{}{+1}{}
  \memrow{65}{+1}{}{}
}}

% The following neighborhood is dense
\newcommand\mooremap{\memtable{\memxyz}}

% Tables
\begin{table}
\begin{tabular}{llll}
  \hline
  From & To & Type & Contents \\
  \hline
  \hex{0000} & \hex{00FF} & RAM (storage-mapped) & State of cell $(x,y,z)$ \\
  \hex{0100} & \hex{01FF} & RAM (transient) & Stack: initialized to zero \\
  \hex{0200} & \hex{03FF} & RAM (transient) & Scratch memory: initialized to zero \\
  \hex{0400} & \hex{3FFF} & ROM & Reserved: zero \\
  \hex{4000} & \hex{7FFF} & RAM (storage-mapped) & Memory-mapped neighborhood (Table~\ref{tab:MooreNeighborhood}) \\
  \hex{8000} & \hex{FFF0} & ROM & Operating system \\
  \hex{FFF0} & \hex{FFF5} & ROM & Reserved; zero \\
  \hex{FFF6} & \hex{FFF7} & Read-only & \code{TIME}: number of cycles left until termination \\
  \hex{FFF8} & \hex{FFF8} & Write-only & \code{SWAP}: swap storage page with neighbor \\
  \hex{FFF9} & \hex{FFF9} & Read-only & \code{RAND}: random byte generator \\
  \hex{FFFA} & \hex{FFFF} & ROM & Reset and interrupt vectors: zero \\
  \hline
\end{tabular}
\caption{
  \label{tab:MemoryMap}
  Memory map for the 6502 virtual machine of cell $(x,y,z)$.
}
\end{table}

\begin{table}
\mooremap
\caption{
  \label{tab:MooreNeighborhood}
  Memory map for the 6502 virtual machine of cell $(x,y,z)$,
  using the 26-cell three-dimensional Moore neighborhood (Figure~\ref{fig:UnitCube}).
  All address ranges not defined in this table are read-only and zero;
  writing to these addresses has no effect.
  Pages mapping to cells outside the chunk (e.g. $x<L\alpha_x$ or $x \geq (L+1)\alpha_x$, or similarly for $y,z$)
  are read-only and zero (unless otherwise specified, as in Section~\ref{sec:AdjacentCells}).
  Writing to these addresses has no effect.
}
\end{table}

\begin{table}
\vonneumannmap
\caption{
  \label{tab:VonNeumannNeighborhood}
  Memory map for the 6502 virtual machine of cell $(x,y,z)$,
  using the 6-cell three-dimensional von Neumann neighborhood.
  This memory map may be considered a subset of the memory map in Table~\ref{tab:MooreNeighborhood}.
}
\end{table}

@@

\begin{figure}
  \includegraphics[width=\textwidth]{unitcube.pdf}
\caption{
  The unit cube of the Moore neighborhood, and its relationship to the memory map of Table~\ref{tab:MooreNeighborhood}.
  The central cell is at $(1,1,1)$, and is mapped to zero page.
  All other cells are mapped to pages in the range \hexrange{4000}{7FFF},
  the first six bits of the address MSB corresponding to the $x$, $y$ and $z$ co-ordinates within this cell (see Table~\ref{tab:AddressBits}).
  \label{fig:UnitCube}
}
\end{figure}

>>>>

% Memory address
\newcommand\code[1]{{\tt #1}}
\newcommand\boldcode[1]{\code{#1}}
\newcommand\hex[1]{{\tt 0x#1}}
\newcommand\hexrange[2]{\hex{#1}{\tt -}\hex{#2}}

@@

% Tables, macros
% Macro to generate neighborhood memory map table
\newcommand\memtable[1]{
\begin{tabular}{rlll}
  \hline
  Label & From & To & Contents \\
  \hline
  #1
  \hline
\end{tabular}
}

\newcommand\memrow[5]{
    \boldcode{#5} & \hex{{#1}00} & \hex{{#1}FF} & State of cell $(x#2,y#3,z#4)$ \\
}

% The \memxyz macro gives a densely-connected (von Neumann) neighborhood
\newcommand\memz[6]{
  \memrow{#1}{#4}{#5}{-1}{D#6}
  \memrow{#2}{#4}{#5}{}{#6}
  \memrow{#3}{#4}{#5}{+1}{U#6}
}

\newcommand\memyz[3]{
  \memz{{#1}0}{{#1}1}{{#1}2}{#2}{-1}{S#3}
  \memz{{#1}4}{{#1}5}{{#1}6}{#2}{}{#3}
  \memz{{#1}8}{{#1}9}{{#1}A}{#2}{+1}{N#3}
}

\newcommand\memxyz{
  \memyz{4}{-1}{W}

  %  \memyz{5}{} includes zero page-mapped cell (x,y,z), which we want to skip, so spell it out without that
  \memz{50}{51}{52}{}{-1}{S}
  \memrow{54}{}{}{-1}{W}
  \memrow{56}{}{}{+1}{E}
  \memz{58}{59}{5A}{}{+1}{N}

  \memyz{6}{+1}{E}
}

% The following table is a sparse neighborhood (von Neumann, 6 neighbors/cell)
\newcommand\vonneumannmap{\memtable{
  \memrow{45}{-1}{}{}{W}
  \memrow{51}{}{-1}{}{S}
  \memrow{54}{}{}{-1}{D}
  \memrow{56}{}{}{+1}{U}
  \memrow{59}{}{+1}{}{N}
  \memrow{65}{+1}{}{}{E}
}}

% The following neighborhood is dense (Moore, 26 neighbors/cell)
\newcommand\mooremap{\memtable{\memxyz}}

% Captions
\newcommand\labeq[2]{\boldcode{#1}\code{=}\hex{#2}}
\newcommand\asslabels{
  The labels (\code{N}, \code{S}, \code{E}, \code{W}, \code{U}, \code{D} \ldots)
  are pre-defined in the assembler to the respective base addresses (\labeq{N}{5900}, \labeq{S}{5100}, and so on).
  The corresponding labels with \code{PAGE} appended
  are predefined to be the respective MSBs of those base addresses (\labeq{NPAGE}{59}, \labeq{SPAGE}{51}, and so on).
  The assembler also pre-defines the symbols
  \labeq{XDIR}{1000}, \labeq{YDIR}{0400}, \labeq{ZDIR}{0100};
  \labeq{XPAGE}{10}, \labeq{YPAGE}{04}, \labeq{ZPAGE}{01};
  \labeq{ORIG}{5500}, \labeq{ORIGPAGE}{55};
  \labeq{BASE}{4000}, \labeq{BASEPAGE}{40}.
}

% Tables
\begin{table}
\begin{tabular}{llll}
  \hline
  From & To & Type & Contents \\
  \hline
  \hex{0000} & \hex{00FF} & RAM (storage-mapped) & State of cell $(x,y,z)$ \\
  \hex{0100} & \hex{01FF} & RAM (transient) & Stack: initialized to zero \\
  \hex{0200} & \hex{03FF} & RAM (transient) & Scratch memory: initialized to zero \\
  \hex{0400} & \hex{3FFF} & ROM & Reserved: zero \\
  \hex{4000} & \hex{7FFF} & RAM (storage-mapped) & Memory-mapped neighborhood (Table~\ref{tab:MooreNeighborhood}) \\
  \hex{8000} & \hex{FFF0} & ROM & Operating system \\
  \hex{FFF0} & \hex{FFF5} & ROM & Reserved; zero \\
  \hex{FFF6} & \hex{FFF7} & Read-only & \code{TIME}: number of cycles left until termination \\
  \hex{FFF8} & \hex{FFF8} & Write-only & \code{SWAP}: write to swap cell with neighbor \\
  \hex{FFF9} & \hex{FFF9} & Read-only & \code{RAND}: random byte generator \\
  \hex{FFFA} & \hex{FFFF} & ROM & Reset and interrupt vectors: zero \\
  \hline
\end{tabular}
\caption{
  \label{tab:MemoryMap}
  Memory map for the 6502 virtual machine of cell $(x,y,z)$.
}
\end{table}

\begin{table}
\mooremap
\caption{
  \label{tab:MooreNeighborhood}
  Memory map for the 6502 virtual machine of cell $(x,y,z)$,
  using the 26-cell three-dimensional Moore neighborhood (Figure~\ref{fig:UnitCube}).
  All address ranges not defined in this table are read-only and zero;
  writing to these addresses has no effect.
  Pages mapping to cells outside the chunk (e.g. $x<L\alpha_x$ or $x \geq (L+1)\alpha_x$, or similarly for $y,z$)
  are read-only and zero (unless otherwise specified, as in Section~\ref{sec:AdjacentCells}).
  Writing to these addresses has no effect.
  \asslabels
}
\end{table}

\begin{table}
\vonneumannmap
\caption{
  \label{tab:VonNeumannNeighborhood}
  Memory map for the 6502 virtual machine of cell $(x,y,z)$,
  using the 6-cell three-dimensional von Neumann neighborhood.
  This memory map may be considered a subset of the memory map in Table~\ref{tab:MooreNeighborhood}.
  \asslabels
}
\end{table}

@@

\begin{figure}
  \includegraphics[width=\textwidth]{unitcube.pdf}
\caption{
  The memory-mapped neighborhood of a cell (see Table~\ref{tab:MemoryMap} and Table~\ref{tab:MooreNeighborhood}).
  The central cell is at $(1,1,1)$, and is mapped to zero page.
  All other cells are mapped to pages in the range \hexrange{4000}{7FFF},
  the first six bits of the address MSB corresponding to the $x$, $y$ and $z$ co-ordinates within this cell (see Table~\ref{tab:AddressBits}).
  \asslabels
  \label{fig:UnitCube}
}
\end{figure}
